---
title: "Iris"
output: html_document
editor_options: 
chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Iris data
```{r}
library(datasets)
iris <- read.csv("C:\\Users\\study\\Downloads\\iris.txt")
head(iris)
```

## Attach the iris dataset
```{r}
attach(iris)
```

## Five number Summary of the numerical variables
```{r}
summary(iris)
```

## Sum, Mean and standartd deviation of Petal.Length

```{r}
sum(p.len)
mean(p.len)
sd(p.len)
```

## Find the sum, mean, median, and standard deviation of s.wid.
```{r}

```




## Sum, Mean and standartd deviation of Sepal.Length, Sepal.Withd
```{r}
apply(iris[,1:2], 2, sum)
apply(iris[,1:2], 2, mean)
apply(iris[,1:2], 2, sd)
```
## Find Sum, Mean and Standartd Deviation of Petal.Length, and  Petal.Width
```{r}

```
## Sum, Mean and standartd deviation of Petal.Length by class

```{r}
aggregate(p.len, by=list(class), FUN=mean)
aggregate(p.len~class, FUN=mean)
aggregate(p.len, by=list(class), FUN=sd)
aggregate(p.len~class, FUN=sd)

```


## Find Sum, Mean and standartd deviation of Width.Length by class

```{r}


```



## Use 'dplyr'
```{r}
library(dplyr)

iris %>%
  group_by(class) %>%
  summarise(sum=sum(p.len), mean=mean(p.len), std=sd(p.len))

iris %>%
  group_by(class) %>%
  summarise(across(c(p.len,s.len),sum))
```

## Use 'dplyr' to get median, min, and max of p.wid
## Use 'dplyr' to get median of p.wid and s.wid
```{r}



```



## Scatterplot Sepal.Length vs Petal.Length
```{r}
plot(s.len, p.len)
```

## Scatterplot Sepal.Length vs Petal.Length using a specific symbol
```{r}
plot(s.len, p.len, pch=18)
```

## Make a scatterplot for s.len and p.len with points of different colors by Species (class)
```{r}
class <- factor(class)
plot(s.len, p.len,col=class, pch=19)
```

## Changing the axis labels, adding a title
```{r}
plot(s.len, p.len,
     col=class, 
     pch=19,
     main="Scatterplot: length of Sepal vs  Petal",
     xlab="length of Sepal",
     ylab="length of Petal")
```

## Make a scatterplot for s.wid and p.wid with points of different colors by Species (class) with a symbol: solid triangle. Add the axis labels, adding a title.
```{r}


```









## Make a histogram for Sepal.Width (red, break points=10).
```{r}
hist(s.wid, breaks=10)
```


## Make overlaying histograms for  
## Sepal.Width:  color is red, break points is 10 and 
## Sepal.Length: color is yellow, break points 20.
## Change labels and title appropriately

```{r}
hist(s.wid, col="red", breaks=10, xlim=c(1,8))
hist(s.len,col="yellow", breaks=20, add=TRUE)
```

## Make boxplots for Sepal.Width, Sepal.Length by species with Title
## (assign colors red for setosa, yellow for versicolor and blue for virginica)
```{r}
boxplot(s.len ~ class, col=c("red", "yellow","blue"))
boxplot(s.wid ~ class, col=c("red", "yellow","blue"))
```










## ggplot
```{r}
library(ggplot2)
ggplot(data=iris, aes(x=s.len, y=p.len))+
       geom_point()
```


## Make a scatterplot Sepal.Length vs Petal.Length with best fit line
```{r}
library(ggplot2)
ggplot(data=iris, aes(x=s.len, y=p.len))+
  geom_point() +                       
  geom_smooth(method="lm")
```

## Make a scatterplot with bestSepal.Length vs Petal.Length
## with best fit line by species

```{r}
ggplot(iris, aes(x=s.len, y=p.len, color=class, fill=class)) +
       geom_point() +
       geom_smooth(method="lm")
```

## Sshow three best fit lines on the three separated panel
```{r}
ggplot(iris, aes(x=s.len, y=p.len, color=class, fill=class)) +
       geom_point() +
       geom_smooth(method="lm") +
       facet_wrap(~class)
```



## Make Boxplots of Sepal.Width by Species with labels and title

```{r}
 ggplot(data=iris, aes(x=class, y=s.wid)) + 
        geom_boxplot()
```
```{r}
 ggplot(data=iris, aes(x=class, y=s.wid)) + 
        geom_boxplot()+
        geom_jitter(width=0.05)
```

```{r}
 ggplot(data=iris, aes(x=class, y=s.wid, fill=class)) + 
        geom_boxplot()+
        geom_jitter(width=0.05)
```


```{r}
ggplot(data=iris, aes(x=class, y=s.wid, fill=class))+
       geom_boxplot() +
       geom_jitter(width=.05) +
       theme_bw() +
       xlab("Class") +
       ylab("Sepal Width") +
       ggtitle("Plot of Width of Sepal by Class")
```



## Make Boxplots of length of petal by class with labels and title. jitter with 0.1, theme_classic, change the name of axes.

```{r}


```